Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2017-03-08 manuscript/supplemental_material.Rnw
switch away from kegg pathview to KGML2Graph

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=15pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage[macros=false]{acro}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature,citestyle=numeric-comp,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=blue,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\usepackage{xspace}
% \usepackage[below]{placeins}
% \makeatletter
% \AtBeginDocument{%
%   \expandafter\renewcommand\expandafter\subsection\expandafter
%     {\expandafter\@fb@secFB\subsection}%
%   \newcommand\@fb@secFB{\FloatBarrier
%     \gdef\@fb@afterHHook{\@fb@topbarrier \gdef\@fb@afterHHook{}}}%
%   \g@addto@macro\@afterheading{\@fb@afterHHook}%
%   \gdef\@fb@afterHHook{}%
% }
% \makeatother
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.6\\textheight",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE,cache.path="supplemental_cache/")
opts_chunk$set(fig.path="supplemental_figure/")
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -2, digits = 2)
library("data.table")
library("ggplot2")
library("reshape2")
library("xtable")
library("grid")
library("scales")
library("gridExtra")
library("ape")
library("ggtree")
library("RSQLite")
library("topGO")
library("org.Hs.eg.db")
library("magrittr")
library("tidyr")

@@

\begin{table}[hb]
  \footnotesize
<<lau.lineage.specific,echo=FALSE,results="asis">>=
lau.all.lin.spec <- placenta.lineage.specific.table("lau.all",fdr.cutoff=0.5)
setkey(lau.all.lin.spec,"Human Symbol")
@ 
\caption{\lineagetableexplanation{the laurasiatheria
    superorder}{all other species}}
\label{tab:laurasiatheria_lineage_specific}
\end{table}

@@

\begin{figure}[tbh]
<<core.placenta.transcriptome.housekeeping,fig.width=6,fig.height=8,out.height="0.8\\textheight",echo=FALSE>>=
cptlh <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species) & housekeeping==TRUE,]
cptlh <- cptlh[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]

print(ggplot(cptlh[FPKM > 0.9,],aes(x=human_name,y=FPKM))+
      geom_boxplot(outlier.shape=NA)+
      scale_y_log10(labels=trans_format("log10", math_format(10^.x)))+
      geom_point(aes(color=Species,shape=Species))+
      scale_shape_manual("Species",
                         guide=guide_legend(#nrow=1,
                                            keywidth=unit(3,"mm"),
                                            keyheight=unit(3,"mm"),
                                            label.theme=element_text(face="italic",
                                                                     angle=0,
                                                                     size=9),
                                            title.position="top"
                                            ),
                         values=0:18)+
      # scale_color_discrete(NULL)+
      coord_flip()+
      xlab("Human Ortholog Name")+
      ylab("Gene Expression (FPKM)")+
      theme_bw() +
      theme(legend.position=c(0.7,0.15),
            axis.text.y=element_text(face="italic",size=5),
            legend.spacing=unit(0,"mm"),
            legend.justification="left",
            plot.margin=unit(c(0,0,0,0),"mm")
            ))
@ 
\caption[Housekeeping genes excluded from the Core Placenta Transcriptome]{
  Housekeeping genes excluded from the Core Placenta Transcriptome.
  \Sexpr{cptlh[,length(unique(human_name))]}
  1:1 \human ortholog genes expressed ($\mathrm{FPKM} \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied which are housekeeping genes~%
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  ordered (from top to bottom) by median expression.}
\label{fig:core_placenta_transcriptome_housekeeping}
\end{figure}

@@

      + geom_segment(mapping=aes(y=V1-ifelse(V1<160,40,0),

@@

                      y=V1-ifelse(V1<160,40,0),

@@

\begin{figure}[tbh]
<<core_placenta_tree_transcriptome,fig.width=7,fig.height=8,out.height="0.8\\textheight",echo=FALSE>>=
cpttl <-
    core.placenta.tree.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                                 Species) & housekeeping==FALSE,]
cpttl <- cpttl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                                levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                            capfirst(names(species.names))))]

print(ggplot(cpttl[FPKM > 0.1,],aes(x=tree_genes,y=FPKM))+
      geom_boxplot(outlier.shape=NA)+
      scale_y_log10(labels=trans_format("log10", math_format(10^.x)))+
      geom_point(aes(color=Species,shape=Species))+
      scale_shape_manual("Species",
                         guide=guide_legend(#nrow=1,
                                            keywidth=unit(3,"mm"),
                                            keyheight=unit(3,"mm"),
                                            label.theme=element_text(face="italic",
                                                                     angle=0,
                                                                     size=9),
                                            title.position="top"
                                            ),
                         values=0:18)+
      # scale_color_discrete(NULL)+
      coord_flip()+
      xlab("Genes in Tree")+
      ylab("Gene Expression (FPKM)")+
      theme_bw() +
      theme(legend.position=c(0.7,0.15),
            axis.text.y=element_text(face="italic",size=5),
            legend.spacing=unit(0,"mm"),
            legend.justification="left",
            plot.margin=unit(c(0,0,0,0),"mm")
            ))
@ 
\caption[Tree-based Core Placenta Transcriptome]{
  Tree-based Core Placenta Transcriptome.
  \Sexpr{cpttl[,length(unique(tree_genes))]}
  gene trees with mean expression ($\mathrm{FPKM} \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied which are not housekeeping~%
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  ordered (from top to bottom) by median expression.}
\label{fig:core_placenta_tree_transcriptome}
\end{figure}

\begin{figure}
<<core_placenta_tree_transcriptome_housekeeping,fig.width=9,fig.height=18,out.height="0.8\\textheight",echo=FALSE>>=
cpttlh <-
    core.placenta.tree.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                                 Species) & housekeeping==TRUE,]
cpttlh <- cpttlh[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                                levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                            capfirst(names(species.names))))]

print(ggplot(cpttlh[FPKM > 0.1,],aes(x=tree_genes,y=FPKM))+
      geom_boxplot(outlier.shape=NA)+
      scale_y_log10(labels=trans_format("log10", math_format(10^.x)))+
      geom_point(aes(color=Species,shape=Species))+
      scale_shape_manual("Species",
                         guide=guide_legend(#nrow=1,
                                            keywidth=unit(3,"mm"),
                                            keyheight=unit(3,"mm"),
                                            label.theme=element_text(face="italic",
                                                                     angle=0,
                                                                     size=9),
                                            title.position="top"
                                            ),
                         values=0:18)+
      # scale_color_discrete(NULL)+
      coord_flip()+
      xlab("Genes in Tree")+
      ylab("Gene Expression (FPKM)")+
      theme_bw() +
      theme(legend.position=c(0.7,0.15),
            axis.text.y=element_text(face="italic",size=5),
            legend.spacing=unit(0,"mm"),
            legend.justification="left",
            plot.margin=unit(c(0,0,0,0),"mm")
            ))
@ 
\caption[Tree-based Core Placenta Transcriptome: Housekeeping Genes]{
  Tree-based Core Placenta Transcriptome: Housekeeping Genes.
  \Sexpr{cpttlh[,length(unique(tree_genes))]}
  gene trees with mean expression ($\mathrm{FPKM} \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied which are housekeeping~%
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  ordered (from top to bottom) by median expression.}
\label{fig:core_placenta_tree_transcriptome_housekeeping}
\end{figure}

@@

\begin{figure}
<<egf_pathway,echo=FALSE,message=FALSE,fig.width=6,fig.height=6,fig.out.width="\\textwidth">>=
plot_pathway("hsa04012")
@
\caption{EGF KEGG pathway}
\label{fig:egf_kegg_pathway}
\end{figure}

<<actin_pathway,echo=FALSE,results="hide",message=FALSE>>=
library("pathview")
pathview(gene.data=plac_path_egid(quote(as.numeric(median_expression))),
         pathway.id="hsa04810",
         both.dirs=list(gene=FALSE,cpd=FALSE),
         limit=list(gene=200,cpd=29),
         high=list(gene="blue"),
         net.signature=FALSE)
@

\begin{figure}
  \includegraphics{hsa04810.pathview.png}
  \caption{KEGG Actin Cytoskeleton Pathway}
  \label{fig:kegg_actin}
\end{figure}

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=15pt]{geometry}
\usepackage{url}
\usepackage{siunitx}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{booktabs}
%\usepackage{float}
\usepackage{multirow}
\usepackage{multicol}
\usepackage[macros=false]{acro}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage[export]{adjustbox}[2011/08/13]
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature,citestyle=numeric-comp,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=blue,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\usepackage{xspace}
% \usepackage[below]{placeins}
% \makeatletter
% \AtBeginDocument{%
%   \expandafter\renewcommand\expandafter\subsection\expandafter
%     {\expandafter\@fb@secFB\subsection}%
%   \newcommand\@fb@secFB{\FloatBarrier
%     \gdef\@fb@afterHHook{\@fb@topbarrier \gdef\@fb@afterHHook{}}}%
%   \g@addto@macro\@afterheading{\@fb@afterHHook}%
%   \gdef\@fb@afterHHook{}%
% }
% \makeatother
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.6\\textheight",out.extra="keepaspectratio,center")
opts_chunk$set(cache=TRUE, autodep=TRUE,cache.path="supplemental_cache/")
opts_chunk$set(fig.path="supplemental_figure/")
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -2, digits = 2)
library("data.table")
library("ggplot2")
library("reshape2")
library("xtable")
library("grid")
library("scales")
library("gridExtra")
library("ape")
library("ggtree")
library("RSQLite")
library("topGO")
library("org.Hs.eg.db")
library("magrittr")
library("tidyr")

@@

\begin{table}[hb]
  \footnotesize
<<lau.lineage.specific,echo=FALSE,results="asis">>=
lau.all.lin.spec <- placenta.lineage.specific.table("lau.all",fdr.cutoff=1)
setkey(lau.all.lin.spec,"Human Symbol")
@ 
\caption{\lineagetableexplanation{the laurasiatheria
    superorder}{all other species}}
\label{tab:laurasiatheria_lineage_specific}
\end{table}

@@

\begin{figure}[tbh]
<<core.placenta.transcriptome.housekeeping,fig.width=6,fig.height=8,out.height="0.8\\textheight",echo=FALSE>>=
cptlh <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species) & housekeeping==TRUE,]
cptlh <- cptlh[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]
cptlh <- cptlh[,median_FPKM:=median(FPKM),by=human_name]
cptlh <- cptlh[,facet:="high"]
cptlh <- cptlh[median_FPKM < median(median_FPKM),facet:="low"]
print(ggplot(cptlh[FPKM > 0.9,],aes(x=human_name,y=FPKM))+
      geom_boxplot(outlier.shape=NA)+
      scale_y_log10(breaks=c(10^1,10^2,10^3,10^4),
                    labels=trans_format("log10", math_format(10^.x)))+
      geom_point(aes(color=Species,shape=Species))+
      scale_shape_manual(NULL,##"Species",
                         guide=guide_legend(#nrow=1,
                                            keywidth=unit(3,"mm"),
                                            keyheight=unit(3,"mm"),
                                            label.theme=element_text(face="italic",
                                                                     angle=0,
                                                                     size=7),
                                            title.position="top"
                                            ),
                         values=0:18)+
      scale_color_discrete(NULL)+
      coord_flip()+
      xlab("Human Ortholog Name")+
      ylab("Gene Expression (FPKM)")+
      theme_bw() +
      expand_limits(y=c(10^4)) + 
      theme(text=element_text(size=7.5),
            legend.position=c(0.99,0.01),
            axis.text.y=element_text(face="italic",size=7.5),
            legend.margin=margin(0,0,0,0,unit="mm"),
            legend.spacing=unit(0,"mm"),
            legend.justification=c(1,0),
            plot.margin=unit(c(0,0,0,0),"mm"),
            strip.background = element_blank(),
            strip.text.x = element_blank()
            )
      + facet_wrap(~facet,scales="free")
      )
@ 
\caption[Housekeeping genes excluded from the Core Placenta Transcriptome]{
  Housekeeping genes excluded from the Core Placenta Transcriptome.
  \Sexpr{cptlh[,length(unique(human_name))]}
  1:1 \human ortholog genes expressed ($\mathrm{FPKM} \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied which are housekeeping genes~%
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  ordered (from top to bottom) by median expression.}
\label{fig:core_placenta_transcriptome_housekeeping}
\end{figure}

@@

      + geom_segment(mapping=aes(y=V1-ifelse(V1<170,40,0),

@@

                      y=V1-ifelse(V1<170,40,0),

@@

\begin{figure}[tbh]
<<core_placenta_tree_transcriptome,fig.width=12,fig.height=8,out.width="\\textwidth",out.height="0.8\\textheight",echo=FALSE>>=
cpttl <-
    core.placenta.tree.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                                 Species) & housekeeping==FALSE,]
cpttl <- cpttl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                                levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                            capfirst(names(species.names))))]
cpttl <- cpttl[,median_FPKM:=median(FPKM,na.rm=TRUE),by=tree_genes]
cpttl <- cpttl[,facet:="high"]
cpttl <- cpttl[median_FPKM < median(median_FPKM),facet:="low"]
print(ggplot(cpttl[FPKM > 0.1,],aes(x=reorder(tree_genes,median_FPKM),y=FPKM))+
      geom_boxplot(outlier.shape=NA)+
      scale_y_log10(breaks=c(10^1,10^2,10^3,10^4),
                    labels=trans_format("log10", math_format(10^.x)))+
      geom_point(aes(color=Species,shape=Species))+
      scale_shape_manual(NULL,##"Species",
                         guide=guide_legend(#nrow=1,
                                            keywidth=unit(3,"mm"),
                                            keyheight=unit(3,"mm"),
                                            label.theme=element_text(face="italic",
                                                                     angle=0,
                                                                     size=7),
                                            title.position="top"
                                            ),
                         values=0:18)+
      scale_color_discrete(NULL)+
      coord_flip()+
      xlab("Genes in Tree")+
      ylab("Gene Expression (FPKM)")+
      theme_bw() +
      expand_limits(y=c(10^4)) + 
      theme(text=element_text(size=7.5),
            legend.position=c(0.99,0.01),
            axis.text.y=element_text(face="italic",size=7.5),
            legend.margin=margin(0,0,0,0,unit="mm"),
            legend.spacing=unit(0,"mm"),
            legend.justification=c(1,0),
            plot.margin=unit(c(0,0,0,0),"mm"),
            strip.background = element_blank(),
            strip.text.x = element_blank()
            )
      + facet_wrap(~facet,scales="free")
      )
@ 
\caption[Tree-based Core Placenta Transcriptome]{
  Tree-based Core Placenta Transcriptome.
  \Sexpr{cpttl[,length(unique(tree_genes))]}
  gene trees with mean expression ($\mathrm{FPKM} \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied which are not housekeeping~%
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  ordered (from top to bottom) by median expression.}
\label{fig:core_placenta_tree_transcriptome}
\end{figure}

\begin{figure}
<<core_placenta_tree_transcriptome_housekeeping,fig.width=12,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
cpttlh <-
    core.placenta.tree.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                                 Species) & housekeeping==TRUE,]
cpttlh <- cpttlh[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                                levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                            capfirst(names(species.names))))]

cpttlh <- cpttlh[,median_FPKM:=median(FPKM,na.rm=TRUE),by=tree_genes]
cpttlh <- cpttlh[,facet:="high"]
cpttlh <- cpttlh[median_FPKM < median(median_FPKM),facet:="low"]
print(ggplot(cpttlh[FPKM > 0.1,],aes(x=reorder(tree_genes,median_FPKM),y=FPKM))+
      geom_boxplot(outlier.shape=NA)+
      scale_y_log10(breaks=c(10^1,10^2,10^3,10^4),
                    labels=trans_format("log10", math_format(10^.x)))+
      geom_point(aes(color=Species,shape=Species))+
      scale_shape_manual(NULL,##"Species",
                         guide=guide_legend(#nrow=1,
                                            keywidth=unit(3,"mm"),
                                            keyheight=unit(3,"mm"),
                                            label.theme=element_text(face="italic",
                                                                     angle=0,
                                                                     size=7),
                                            title.position="top"
                                            ),
                         values=0:18)+
      scale_color_discrete(NULL)+
      coord_flip()+
      xlab("Genes in Tree")+
      ylab("Gene Expression (FPKM)")+
      theme_bw() +
      expand_limits(y=c(10^4)) + 
      theme(text=element_text(size=7.5),
            legend.position=c(0.99,0.01),
            axis.text.y=element_text(face="italic",size=7.5),
            legend.margin=margin(0,0,0,0,unit="mm"),
            legend.spacing=unit(0,"mm"),
            legend.justification=c(1,0),
            plot.margin=unit(c(0,0,0,0),"mm"),
            strip.background = element_blank(),
            strip.text.x = element_blank()
            )
      + facet_wrap(~facet,scales="free")
      )
@ 
\caption[Tree-based Core Placenta Transcriptome: Housekeeping Genes]{
  Tree-based Core Placenta Transcriptome: Housekeeping Genes.
  \Sexpr{cpttlh[,length(unique(tree_genes))]}
  gene trees with mean expression ($\mathrm{FPKM} \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied which are housekeeping~%
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  ordered (from top to bottom) by median expression.}
\label{fig:core_placenta_tree_transcriptome_housekeeping}
\end{figure}

@@

\begin{figure}
<<egf_pathway,echo=FALSE,message=FALSE,fig.width=6,fig.height=6,fig.out.width="\\textwidth">>=
plot_pathway("hsa04012")
@
\caption{EGF KEGG pathway.}
\label{fig:egf_kegg_pathway}
\end{figure}

\begin{figure}
<<actin_pathway,echo=FALSE,results="hide",message=FALSE,warning=FALSE>>=
plot_pathway("hsa04810",pathcolor=quote(log10(median_expression+1)),cutpoint=0.1)
#library("pathview")
# pathview(gene.data=plac_path_egid(quote(as.numeric(median_expression))),
#          pathway.id="hsa04810",
#          both.dirs=list(gene=FALSE,cpd=FALSE),
#          limit=list(gene=200,cpd=29),
#          high=list(gene="blue"),
#          net.signature=FALSE)
@

%  \includegraphics{hsa04810.pathview.png}
  \caption{KEGG Actin Cytoskeleton Pathway. Expression in $\log_{10}$ FPKM.}
  \label{fig:kegg_actin}
\end{figure}
